---
title: "Homework 2"
author: "Jessica Flynn"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1 

Read the Mr. Trashwheel dataset. 
```{r trashwheel_df}
trashwheel_df = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate( 
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read in precipitation data for 2018 and 2017
```{r precip_df_2017_2018}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation", 
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation", 
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)

```


Combine annual precipitation
```{r combine_precip}
precip_df = 
  bind_rows(precip_2017, precip_2018)

``` 

Create month name 
```{r month_name}

month_df = 
  tibble( 
    month = 1:12, 
    month_name = month.name
  )

precip_df = 
  left_join(precip_df, month_df, by = "month")

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, MD. As trash enters the harbor, the trashwheel collects that trash and stores it in a dumpster. The dataset contains information on year, month and trash collected. Some specific trash types that are reported on include `cigarette_butts`, `chip_bags` and `plastic_bottles`, among others. There are also some metrics regarding the trash including `weight_tons` and `volume_cubic_yards`. There are a total of `r nrow(trashwheel_df)` rows in the final dataset and a total of `r ncol(trashwheel_df)` columns. The median number of sportsballs in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.

Additional data sheets include precipitation data by month. Here, we looked specifically at the years 2017 and 2018. There are a total of `r nrow(precip_df)` rows in the data for these two years. The columns in the datset include `r colnames(precip_df)`. The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()`and the total in 2018 was  `r precip_df %>% filter(year == 2017) %>% pull(total) %>% sum()`.

## Problem 2 

## Problem 3 

